A python toolbox to tackle the curse of imbalanced datasets in machine learning

### 2016-06-18 main.tex

% Heading arguments are {volume}{year}{pages}{submitted}{published}{author-full-names}
\jmlrheading{7}{2016}{1-5}{7/16}{-}{Guillaume Lema\^itre, Fernando Nogueira, and Davyid Victor}

% Short headings should be running head and authors last names
\ShortHeadings{A python toolbox to tackle the curse of imbalanced dataset in machine learning}{Lema\^itre, Nogueira, and Victor}
\firstpageno{1}

@@

\title{A python toolbox to tackle the curse of imbalanced dataset in machine learning}
\author{Guillaume Lema\^itre \email g.lemaitre58@gmail.com \\ \addr{LE2I UMR6306, CNRS, Arts et M\'etiers, Universit\'e Bourgogne Franche-Comt\'e} \\ \addr{12 rue de la Fonderie, 71200 Le Creusot, France} \\ \addr{ViCOROB, Universitat de Girona} \\ \addr{Campus Montilivi, Edifici P4, 17071 Girona, Spain}
        \AND
        Fernando Nogueira \email fmfnogueira@gmail.com \\ \addr{theScore, Inc.} \\ \addr{500 King Street West 4\textsuperscript{th} Floor Toronto, Ontario M5V1L9 Canada}
        \AND
        Dayvid Victor Rodrigues de Oliveira \email victor.dvro@gmail.com \\ \addr{VIISAR Research Group, Centto de Inform\'atica (CIn) - Federal University of Pernambuco (UFPE)} \\ \addr{Av. Prof. Luiz Freire, s/n - Cidade Universitária Recife - PE, 50740-540, Brazil}} 
\editor{-}

@@

This problem is frequently referred as ``class imbalance'' problem~\cite{prati2009data} and has been encountered in multiple areas such as telecommunication managements, bioinformatics, fraud detection, and medical diagnosis. 
Imbalanced data substantially compromises the learning process since most of the standard machine learning algorithms expect balanced class distribution or an equal mis-classification cost~\cite{he2009learning}.
Medical data are prone to such drawbacks due to the fact that the portion of diseased samples or patients is far lower than healthy cases.
Furthermore, the detection and classification of minority malignant cases are highly essential so that the sensitivity of developed algorithms need to be maximized.
Consequently, the problem of imbalanced data is usually addressed by employing different techniques which do not impair the topology of the data.

@@

Three strategies can be employed to overcome the problem of imbalanced dataset: (i) under-sampling, (ii) over-sampling, (iii) a combination of both, and (iv) ensemble learning.
The following sections give an overview of the techniques used to tackle this issue.

>>>>

% Heading arguments are {volume}{year}{pages}{submitted}{published}{author-full-names}
\jmlrheading{7}{2016}{1-5}{7/16}{-}{Guillaume Lema\^itre, Fernando Nogueira, and Dayvid V. R. Oliveira}

% Short headings should be running head and authors last names
\ShortHeadings{A python toolbox to tackle the curse of imbalanced datasets in machine learning}{Lema\^itre, Nogueira, and Victor}
\firstpageno{1}

@@

\title{A python toolbox to tackle the curse of imbalanced datasets in machine learning}
\author{Guillaume Lema\^itre \email g.lemaitre58@gmail.com \\ 
    \addr{LE2I UMR6306, CNRS, Arts et M\'etiers, Universit\'e Bourgogne Franche-Comt\'e} \\ 
    \addr{12 rue de la Fonderie, 71200 Le Creusot, France} \\ 
    \addr{ViCOROB, Universitat de Girona} \\ 
    \addr{Campus Montilivi, Edifici P4, 17071 Girona, Spain}
        \AND
        Fernando Nogueira \email fmfnogueira@gmail.com \\ 
        \addr{theScore, Inc.} \\ 
        \addr{500 King Street West 4\textsuperscript{th} Floor Toronto, Ontario M5V1L9 Canada}
        \AND
        Dayvid V. R. Oliveira \email dvro@cin.ufpe.br \\ 
        \addr{VIISAR Research Group, Centro de Inform\'atica - Universidade Federal de Pernambuco} \\ 
        %\addr{Av. Prof. Luiz Freire, s/n - Cidade Universitária Recife - PE, 50740-540, Brazil}} 
        \addr{Av. Jornalista Aníbal Fernandes, s/n - Cidade Universitária - PE, 50740-560, Brazil}} 
\editor{-}

@@

Many real world datasets have many samples of some classes (majority classes),
and only a few samples the other class (minority classes). This imbalance gives
rise to the ``class imbalance'' problem~\cite{prati2009data} (or ``curse of imbalanced datasets'')
which is the problem of learning a concept from the class that has a small number of samples compared
to the other classes.

The class imbalance problem has been encountered in multiple areas such as 
telecommunication managements, bioinformatics, fraud detection, and medical diagnosis,
and has been considered one of the top 10 problems in data mining and 
pattern recognition~\cite{10problemsInDM:2006}. 
Medical data, for example, are prone to suffer from class imbalance due to the fact 
that the portion of diseased patients is far lower than healthy patients,
and the detection and classification of minority diseased patients are highly essential
so that they can be treated as soon as possible.

Imbalanced data substantially compromises the learning process, since most of the 
standard machine learning algorithms expect balanced class distribution or an 
equal misclassification cost~\cite{he2009learning}. For this reason, several
approaches have been specifically proposed to handle such datasets.

In this paper, we present the \textbf{UnbalancedDataset} API, 
\textit{a python toolbox to tackle the curse of imbalanced datasets
in machine learning}. The following sections present the implemented
methods, implemented design, sustainability and continuous integration details,
and finally, the conclusion of this paper, including future functionalities
for the UnbalancedDataset API.

%This problem is frequently referred as ``class imbalance'' problem~\cite{prati2009data} and has been encountered in multiple areas such as telecommunication managements, bioinformatics, fraud detection, and medical diagnosis. 
%Imbalanced data substantially compromises the learning process since most of the standard machine learning algorithms expect balanced class distribution or an equal mis-classification cost~\cite{he2009learning}.
%Medical data are prone to such drawbacks due to the fact that the portion of diseased samples or patients is far lower than healthy cases.
%Furthermore, the detection and classification of minority malignant cases are highly essential so that the sensitivity of developed algorithms need to be maximized.
%Consequently, the problem of imbalanced data is usually addressed by employing different techniques which do not impair the topology of the data.

@@

Three strategies can be employed to overcome the problem of imbalanced datasets: 
(i) under-sampling, (ii) over-sampling, (iii) a combination of both, and (iv) ensemble learning.
The following sections give an overview of the techniques used to tackle this issue.
